Dissecting the relationship between protein structure and sequence variation

### 2015-05-20 manuscript/voronoi/mnv.tex
MN update

\vfill
%\Title{Predicting Sequence Variability from Voronoi Tessellation of Proteins}
\Title{Dissecting the relationship between protein structure and sequence variation}
\vfill
\Author{Amir Shahmoradi$^{1}$, Claus O. Wilke$^2$}
\Address{\affiliation}
\vfill
\begin{Abstract}
    What are the best structural predictors of protein's sequence evolution? A number of site-specific structural characteristics have been proposed over the past decade to answer this question. Most importantly, the role of local packing density has been highlighted and shown recently to be the dominant factor in shaping the observed patterns of site-specific sequence variability of proteins. A wide variety of local packing density measures has been introduced in the literature such as the Weighted Contact Number which represents by definition, the combined effects of local packing density and long-range amino acid interactions on sequence variability. Here we propose a methodology to segregate and quantify the extent to which long-range interactions determine the general patterns of sequence variability in proteins, independently of local packing density. We use Voronoi partitioning of protein's 3-dimensional structure to obtain a parameter-free measure of the local packing density, defined as the inverse volume of the Voronoi cell corresponding to each site in protein.

@@

    In order to assess the prediction power of site-specific variables derived from Voronoi tessellation, first the geometric centers of all side-chains for each of the proteins in dataset were calculated and used as the seeds of Voronoi polyhedra. We show in Appendix \ref{app:best_voronoi} that the choice of the geometric center of side chains as Voronoi seeds -- in contrast to other sets of atomic coordinates representative of individual sites in protein -- results in the strongest correlations of Voronoi cell properties with other site-specific structural and evolutionary characteristics of proteins.  \\

@@

    \subsection*{Average Side-Chain Coordinates as the Best Representation of Protein 3D Structure}
    \label{sec:voronoi_SC}

@@

    \subsection*{Average Side-Chain B Factors as the Best Representation of Local Fluctuations of Amino Acids in Proteins}

@@

>>>>

\vfill
%\Title{Predicting Sequence Variability from Voronoi Tessellation of Proteins}
\Title{Dissecting the relationship between protein structure and sequence variation}
\vfill
\Author{Amir Shahmoradi$^{1}$, Claus O. Wilke$^2$}
\Address{\affiliation}
\vfill
\begin{Abstract}

    What are the best structural predictors of protein's sequence evolution? A number of site-specific structural characteristics have been proposed over the past decade to answer this question. Most importantly, the role of local packing density has been highlighted and shown recently to be the dominant factor in shaping the observed patterns of site-specific sequence variability of proteins. A wide variety of local packing density measures has been introduced in the literature such as the Weighted Contact Number which represents by definition, the combined effects of local packing density and long-range amino acid interactions on sequence variability. Here we propose a methodology to segregate and quantify the extent to which long-range interactions determine the general patterns of sequence variability in proteins, independently of local packing density. We use Voronoi partitioning of protein's 3-dimensional structure to obtain a parameter-free measure of the local packing density, defined as the inverse volume of the Voronoi cell corresponding to each site in protein.

@@

    In order to assess the prediction power of site-specific variables derived from Voronoi tessellation, first the geometric centers of all side-chains for each of the proteins in dataset were calculated and used as the seeds of Voronoi polyhedra. We show in Appendix \ref{app:best_crd} that the choice of the geometric center of side chains as Voronoi seeds -- in contrast to other sets of atomic coordinates representative of individual sites in protein -- results in the strongest correlations of Voronoi cell properties with other site-specific structural and evolutionary characteristics of proteins.  \\

@@

@@

\appendix{Average Side-Chain B Factors as the Best Representation of Local Fluctuations of Amino Acids in Proteins}

@@

    \appendix{Average Side-Chain Coordinates as the Best Representation of Protein 3D Structure}
    \label{app:best_crd}
